Multi-resolution Neural Field Estimation

### 2011-03-04 Multiresolution_framework/ltx/manuscript.tex
minor edit

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, RMSE = 2.10. (b) $j=1$, $n_x=33$, RMSE = 1.42. (c) $j=2$, $n_x=65$, RMSE = 0.80. (d) $j=3$, $n_x=131$, RMSE = 0.72. (e) $j=4$, $n_x=263$, RMSE=0.72.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Kernel.pdf}
		\caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
	\label{fig:KernelEstimate}
\end{figure}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

@@

In order to apply the framework to real data several assumptions must be made, such as stationarity of the parameters over the estimation period. Perhaps the most critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model, which holds for a nonlinear activation function. Furthermore, the theory can be adapted to more sophisticated nonlinear forms of the field equations, where the unscented RTS smoother should be used in the E-step of the estimation algorithm.

>>>>

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 2.10$. (b) $j=1$, $n_x=33$, $RMSE = 1.42$. (c) $j=2$, $n_x=65$, $RMSE = 0.80$. (d) $j=3$, $n_x=131$, $RMSE = 0.72$. (e) $j=4$, $n_x=263$, $RMSE=0.72$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/test.pdf}
		\caption{Connectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Weighted kernel scaling functions at $j=1$. (c) Weighted kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

@@

In order to apply the framework to real data several assumptions must be made, such as stationarity of the parameters over the estimation period. Perhaps the most critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model. The decomposition holds for more sophisticated nonlinear forms of the field equations, where the unscented RTS smoother should be used in the E-step of the estimation algorithm.
